Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-01-09 survey/survey.tex
worked on abstract

>>>>

\begin{abstract}
  Our main models of computation (the Turing Machine and the RAM) and
  most modern computer architectures make fundamental assumptions
  about which primitive operations are realizable on a physical
  computing device. The consensus is that these primitive operations
  include logical operations like conjunction, disjunction and
  negation, as well as reading and writing to a (large) collection to
  memory locations. This perspective conforms to a macro-level view of
  physics and indeed these operations are realizable using macro-level
  devices involving thousands of electrons. This point of view is
  however incompatible with computation realized using quantum devices
  or analyzed using elementary thermodynamics as both these
  fundamental physical theories imply that information is a conserved
  quantity of physical processes and hence of primitive computational
  operations.

  Our aim is re-develop foundational computational models in a way
  that embraces the principle of conservation of information. We first
  define what information is and what its conservation means in a
  computational setting. We emphasize the idea that computations,
  whatever they are, must be reversible transformations on
  data. Thinking topologically, one can think of data as modeled using
  topological spaces and programs as modeled by reversible
  deformations of these spaces. We then illustrate this idea using
  three notions of data and their associated reversible computational
  models. The first instance only assumes unstructured finite data,
  i.e., discrete topological spaces. The corresponding notion of
  reversible computation is that of permutations. We show how this
  simple model subsumes conventional computations on finite sets. We
  then consider a modern structured notion of data based on the
  Curry-Howard correspondence between logic and type theory. We
  develop the corresponding notion of reversible deformations using a
  sound and complete programming language for witnessing type
  isomorphisms and proof terms for commutative semirings.  We then
  ``move up a level'' to examine spaces that treat programs as data
  which is a crucial notion for any universal model of computation. To
  derive the corresponding notion of reversible programs between
  programs, i.e., reversible program equivalences, we look at the
  ``higher dimensional'' analog to commutative semirings: symmetric
  rig groupoids. The coherence laws for these groupoids turn out to be
  exactly the sound and complete reversible program equivalences we
  seek.

  We conclude with some possible generalizations inspired by homotopy
  type theory and survey several open directions for further research.

%%  Something about Theseus?
\end{abstract}
